---
title: "About Me"
description: |
  Some additional details about my work experience
author:
  - name: Margaret Hammond
    url: {https://github.com/mbhammond/a2_hammond_margaret}
date: 2022-02-19
output:
  distill::distill_article:
    self_contained: false
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(purrr)
library(minpack.lm)
library(broom)
library(patchwork)
library(kableExtra)
library(janitor)
library(modelr)
```

# Sustainability & Data Management Experience:

**Sustainable Energy Research Assistant – UCSB Bren School**, Santa Barbara, CA (11/21–Present) 
  - Examined 22 company data center Environmental, Social, and Governmental (ESG) reports, synthesized information regarding SASB, TCFD, GRI standards, reported data in Microsoft Excel Online database.
  
  - Evaluate renewable energy, power use, and water use efficiency for 4 corporations using computer data centers to track the change of green energy efficiency for large scale computing.
  
  - Wrote 4 data center 800-word energy use summaries to be published on a data collection website to educate the public on data center growth and energy efficiencies. 

***GIS Technician – US Geological Survey (USGS)***, Denver, CO (7/19–7/21)
  - Led creation of beryllium and fluorspar US Mineral Deposit GIS data releases comprising of economic and historical data, summarizing data in 500-word abstract detailing critical mineral production and potential to ensure no import reliance in the future.
  
  - Recorded occurrence of unknown mine land from 15 states’ georeferenced topographic maps to help prevent land mismanagement: plotted 50,0000+ mine data entries using Arc GIS 10.8+.
  - Researched beryllium and fluorspar markets, reading 300+ historical and current mining corporation documents and attending 2 Zoom conferences to become USMIN beryllium and fluorspar expert.
  
  - Oversaw 6 employees on the process of correctly recording mine feature points and polygons to create a cohesive database for project posterity.


### Strategic Communications Experience

***Library Engagement Ambassador – University of Michigan Library***, Ann Arbor, MI (9/17–5/19)

  - Publicized and hosted monthly library events attended by 50+ students, including poetry slams, free professional headshots, and magazine workshops to create an inclusive library atmosphere for students. 
  
  - Promoted library services and resources to students with targeted campaigns; created Canva graphics to share with 1800+ followers on Facebook and Twitter 3 times a week.
  

***Explorer Guide – Michigan Department of Natural Resources***, Brighton, MI (5/18–8/18)
  - Designed and led hour-long, twice-daily educational programs at amphitheater for 100+ total campers per day to increase awareness of invasive species, natural resources, wetlands, and other environmental topics. 
  
  - Increased attendance from 20 to 500+ patrons a week from May to August. Publicized events and encouraged attendance by designing Canva flyers, approaching campsites, and creating Facebook posts.
  
# Example of Coding Experience

### 1. Overview
In this task, I will be using the Lightfoot (2020) data on lizards in the Jornada Basin LTER site, New Mexico, 1989-2006. I will be creating a non linear least squares model to estimate the parameters of a length to weight model for the lizard populations. 

### 2. Explore the Data

#### 2.1 Read in the data
```{r}
lizards <- read_csv(here("data", "lizard.csv")) %>% 
  mutate(log_length = log(SV_length), # use log() to linearize the model
         log_weight = log(weight))
```
### 3. Set up the model

Must set up the NLS model of weight as a function of snout to vent length to be able to make guesses for coefficients a and b. Take the data and create a linear model in order for OLS regression to occur.


$$
W = a(SVL)^b
$$
In this model, W is weight, and SVL is snout to vent length.


```{r}
lizards_model <- function(a, SVL, b) {
  W = a*SVL^b
  return(W)
}
```



```{r}
parameter_guess <- lm(log_weight ~ log_length, data = lizards) # Created for OLS

lizards_coef <- coef(parameter_guess) 

# lizards_coef checking the coefficients
```


### 4. Functions
Create the general lizards NLS model of weight as a function of snout to vent length. 

```{r, results = FALSE}

lizards_nls <- nls(weight ~ lizards_model(a, SV_length, b),
                   data = lizards,
                   start = list(
                     a = 2.718^(lizards_coef[1]/lizards_coef[2]),
                     b = lizards_coef[2]),
                   trace = TRUE)
```

#### 4.1 Table of Model
This table shows the estimated parameters for the NLS model. 

```{r}
tidy_nls <- tidy(lizards_nls)

tidy_nls$p.value <- ifelse(tidy_nls$p.value < .001, paste('< .001'))

tidy_nls %>% 
  kable(col.names = c('Parameter', 
                      'Estimated', 
                      'Standard Error', 
                      'Statistic',
                      'P Values'), 
        caption = 'Table 1. Parameter estimates of coefficient a and b for the nls model of weight as a function of snout to vent length of lizards in Jornada Basin LTER, from 1989-2006 (Lightfoot 2020) .',
        digits = 4) %>% 
  kable_classic(bootstrap_options = "striped", full_width = FALSE)
```


#### 4.2 NLS Predictions

I will use the NLS model to generate predictions of weight based on snout to vent length from the Lightfoot, 2020 data. 

```{r}
general_mdl_pred <- predict(lizards_nls)


lizards_complete <- cbind(lizards, general_mdl_pred)

ggplot(data = lizards_complete,
       aes(x = SV_length, 
           y = general_mdl_pred)) +
  geom_point(aes(x = SV_length, y = weight, color = sex)) +
  scale_color_manual(values = c('pink', 'blue2')) +
  geom_line(color = 'indianred2', size = 1) +
  labs(x = 'Snout to Vent Length (mm)',
       y = 'Weight (g)',
       title = "Lizard NLS Model Predictions",
       caption = 'Figure 1. NLS vs measured weight and
       snout to vent length of lizards in
       Jornada Basin LTER site, 1989-2006 (Lightfoot 2020)') +
  theme_minimal()



```


#### Fit NLS to Western Whiptail lizard
I will filter out the dataset for male Western Whiptail lizard (*Cnemidophorus tigrisatus*). Then, I will fit an NLS model to this subset. After, I compare the output from the species specific NLS model to the general NLS model for all species by graphing the model fits on the Western Whiptail male data. 


```{r, results = FALSE}

# filter out just male Western Whiptail

whip_sub <- lizards %>% 
  filter(spp == 'CNTI', 
         sex == 'M') %>% 
  mutate(log_weight = log(weight)) %>% 
  mutate(log_length = log(SV_length))


whip_pred <- lm(log_weight ~ log_length, data = whip_sub)

whip_coef <- coefficients(whip_pred)

whip_nls <- nls(weight ~ lizards_model(a, SV_length, b),
                   data = whip_sub,
                   start = list(
                     a = 2.718^(whip_coef[1]/whip_coef[2]),
                     b = whip_coef[2]),
                   trace = TRUE)

tidy_whip <- tidy(whip_nls)

# tidy_whip

```

#### Compare the Species NLS to General NLS to see which is a best fit for the Western Whiptail species. 

```{r}
lizards_general_whip <- lizards_complete %>% 
  filter(spp == 'CNTI',
         sex == 'M')

whip_mdl_pred <- predict(whip_nls)

complete_pred <- cbind(lizards_general_whip, whip_mdl_pred)

rmse_whip <- rmse(whip_nls, complete_pred)

rmse_gen <- rmse(lizards_nls, complete_pred)

rmse_complete <- cbind(rmse_whip, rmse_gen)
# rmse_complete, rmse_whip has a lower value than rmse_gen, RMSE of whip specific set is better for Western Whiptail species



```

```{r}
# Create plot showing the differences between the general NLS and species specific nls for Western Whip
caption = 'Figure 2. NLS species specific model for Western Whiptail (in red) vs NLS of all lizard species (in navy) in Jornada Basin LTER site, 1989 - 2006. Black points represent actual Western Whiptail measurements of weight (g) to snout to vent length (mm) (Lightfoot 2020). RMSE calculations done for both models, species specific model contained a lower RMSE score of 3.349 compared to general NLS model RMSE of 3.562. The species specific RMSE should be used when doing RMSE calculations for Western Whiptail, it is better fit for the data.'

ggplot(data = complete_pred) +
  geom_point(aes(x = SV_length, y = weight)) +
  geom_line(aes(x = SV_length, y = general_mdl_pred), color = 'red', size = 1) +
  geom_line(aes(x = SV_length, y = whip_mdl_pred), color = 'navy', size = 1) +
  theme_minimal() +
  labs(x = 'Snout to Vent Length (mm)',
       y = 'Weight (g)',
       title = 'General NLS vs. Species Specific NLS',
      caption = str_wrap(caption)) 
```


#### Data Citation
Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f